---
title: "Stability Analyses"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(ggplot2)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r}
# Get the data
tbi_age <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_age.csv')
tbi_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_year.csv')
tbi_military <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_military.csv')

#glimpse(tbi_year)
```


# Explore TBI by category
```{r}
tbi_age %>%
  filter(age_group=='Total') %>%
  group_by(type) %>%
  mutate(rate_est_total = sum(rate_est)) %>%
  ungroup(type) %>%
  mutate(type = stats::reorder(type,rate_est_total)) %>%
  ggplot(aes(x=type, y = rate_est, fill = injury_mechanism))+
  geom_col(width = .6) +
  coord_flip() +theme_bw() +
  theme(legend.title= element_blank(),legend.position = c(.6,.3),
        legend.background=element_rect(color= 'black',linetype='solid')) + 
    labs(x = " ", y="Rate of TBI per 100,000")

```
Most of documented TBIs occur at the hospital, to the point where it does not make sense to analyze these categories separately. 


# How does the rate of TBI increase with age?
```{r}
tbi_age %>%
  # remove age group that encompasses 0-4 and 5-14
  filter(age_group!='0-17' & age_group!='Total') %>%
  # clean up injurt mechanism categories 
  filter(injury_mechanism!="Intentional self-harm") %>% 
  mutate(injury_mechanism = ifelse(grepl("Other", injury_mechanism), "Other", injury_mechanism)) %>%
  mutate(age_group = factor(age_group, levels = c("0-4", "5-14", "15-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75+"))) %>%
  # plot
  ggplot(aes(x=age_group, y = rate_est, fill = injury_mechanism)) + 
  geom_col() +
  labs(x = "Age", y="Rate of TBI per 100,000") + 
  coord_flip() + 
  theme_bw() +
  theme(legend.title= element_blank(),legend.position = c(.75,.5),
        legend.background=element_rect(color= 'black',linetype='solid'),
        text = element_text(size=12))

```
The following age-related trends can be seen:  
- Young people (below 24) are more likely to have a TBI caused by being <b>unintentionally struck or by against an object</b>.  
- <b>Unintentional falls</b> account for more TBIs among toddlers (0-4) and the elderly (75#).  
- <b>Assults</b> account for more TBIs among yound adults (15-24), and adults (25-54)

# Plot trajectory of TBI rates for cases where we saw an age-related trend
```{r}

tbi_age %>%
  # remove age group that encompasses 0-4 and 5-14
  filter(age_group!='0-17' & age_group!='Total') %>%
  # clean up injurt mechanism categories 
  filter(injury_mechanism=="Assault" | injury_mechanism=="Unintentionally struck by or against an object" | injury_mechanism=="Unintentional Falls") %>% 
    mutate(age_group = factor(age_group, levels = c("0-4", "5-14", "15-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75+"))) %>%
  group_by(age_group, injury_mechanism) %>%
  summarize(rate_total = sum(rate_est)) %>%
  ungroup(age_group, injury_mechanism) %>%
  ggplot(aes(x = age_group, y = rate_total, color = injury_mechanism,group=injury_mechanism)) + 
  geom_point() + 
  geom_line() + 
  labs(x = 'Age', y = 'Rate of TBI per 100,000') + 
    theme_bw() +
  theme(legend.title= element_blank(),legend.position = c(.3,.8),
        legend.background=element_rect(color= 'black',linetype='solid'), text = element_text(size=14))
```
The strongest age-related change is the U-shaped trajectory of TBIs due to <b> unintentional falls</b>.
